### 2017-03-09 PaperEx170308.Rmd
paper examples

---
title: "Mixture Modeling Paper Results"
author: "Sarah Sullivan"
date: "March 8, 2017"
output: word_document
---

```{r loadfunctions}
library(mixsmsn)
library(sn)
#-----------------------------------------------
#[IC] Load Functions
#-----------------------------------------------

#-----------------------------------------------
#[IC1]  fitloops
#-----------------------------------------------

#Runs loops of mixture model over different numbers of components and the normal and skew-normal distribution. This function accepts a data frame of the data and identifiers (variables must be entitled data and id respectively). This is the longest function to run and may fail, but can be re-run with adjustments using various options based on the error message output when the function fails. The default values are to do both skewnormal and normal approximation, with up to 5 components and 10 loops per combination of distribution and number of components. It is not reccommended to change these unless the default function fails.
#accepts a dataframe with columns id and data
fitloops<-function(datawithids,loops=10,maxcp=5){
  dists="Both"
  datawithids<-unique(datawithids)
  data<-datawithids$data
  datapts<-length(data)
  norms<-vector("list")
  errors<-matrix(0, nrow=maxcp, ncol=2)
  loopErr = function(data,i,j,family) {
    tryCatch(smsn.mix(data, nu =3, g = i, get.init = TRUE, criteria = TRUE, group = TRUE, family = family, calc.im=FALSE, obs.prob=TRUE, iter.max=5000, kmeans.param=list(iter.max = 50, n.start = 5)),
             error = function(e) {message(paste("Error in loop", j, "of", family, "with", i, "components", sep=" ")); 
               NA}) 
  }

  for (i in 1:maxcp){
    norms$mc[[i]]<-vector("list", maxcp)
    norms$ms[[i]]<-vector("list", maxcp)
    j=1
    while (max(errors)<=2 & j<=loops) {

      norms$mc[[i]][[j]]<-vector("list", 15)
      norms$ms[[i]][[j]]<-vector("list", 15)
      if (dists=="Both") {
        norms$mc[[i]][[j]]<-loopErr(data,i,j,"Normal")  
        norms$ms[[i]][[j]]<-loopErr(data,i,j,"Skew.normal")
      } else if (dists=="Normal"){ 
        norms$mc[[i]][[j]]<-loopErr(data,i,j,"Normal")
      } else if (dists=="Skew.normal"){
        norms$ms[[i]][[j]]<-loopErr(data,i,j,"Skew.normal")
      } else {
        stop(print("dists must be specified as 'Both' or 'Normal' or 'Skew.normal'"))
      }

      message(paste(i, "Component(s), Loop", j))
      if (is.na(norms$mc[[i]][[j]][1])==T) {
        norms$mc[[i]][[j]]<-vector("list")
        norms$mc[[i]][[j]]$bic<-NA
        errors[i,1]<-errors[i,1]+1
      } 
      if (is.na(norms$ms[[i]][[j]][1])==T) {
        norms$ms[[i]][[j]]<-vector("list")
        norms$ms[[i]][[j]]$bic<-NA
        errors[i,2]<-errors[i,2]+1
      } 
      if (dists=="Normal"){ 
        norms$ms[[i]][[j]]<-vector("list")
        norms$ms[[i]][[j]]$bic<-NA
      } else if (dists=="Skew.normal"){
        norms$mc[[i]][[j]]<-vector("list")
        norms$mc[[i]][[j]]$bic<-NA
      }
      j=j+1
      ic=i
    }
  }

  if (i==maxcp & j==loops+1 & max(errors[maxcp, 2])<=2 & max(errors[maxcp, 1])<=2 ) {
    nobj<-vector("list")
    nobj$mc<-matrix(nrow=maxcp, ncol=loops)
    nobj$ms<-matrix(nrow=maxcp, ncol=loops)

    for (i in 1:maxcp){
      for (j in 1:loops){
        nobj$mc[i,j]<-norms$mc[[i]][[j]]$bic
        nobj$ms[i,j]<-norms$ms[[i]][[j]]$bic
      }
    }

    for (i in 1:maxcp){
      if (dists=="Both" | dists=="Normal"){
        nobj$lmin$mc[i]<-which.min(nobj$mc[i,])
      } else {
        nobj$lmin$mc[i]<-1
      }
      if (dists=="Both" | dists=="Skew.normal"){
        nobj$lmin$ms[i]<-which.min(nobj$ms[i,])
      } else {
        nobj$lmin$ms[i]<-1
      }
    }

    fitres<-vector("list")
    for (i in 1:maxcp){
      fitres$no[[i]]<-vector("list", 15)
      fitres$sn[[i]]<-vector("list", 15)
      fitres$no[[i]]<-norms$mc[[i]][[nobj$lmin$mc[i]]]
      fitres$sn[[i]]<-norms$ms[[i]][[nobj$lmin$ms[i]]]
    }
    fitres$loops<-loops
    fitres$maxcp<-maxcp
    fitres$datawithids<-datawithids
    fitres$errors<-errors
    return(fitres)
  } else {
    if (which.max(errors)<=maxcp){
      dist<-"Normal"
    } else {
      dist<-"Skew.normal"
    }
    if (ic<=2){
      stop(paste("More than two loops of the mixture modeling algorithm failed for the", dist, "distribution with", ic, "component(s). First ensure that the necessary packages have been loaded correctly. If the packages are loaded correctly, check for extreme outliers (which may be dropped with extreme caution) or violations of skewnormal or normal distribution assumptions."))
    } else {
      stop(paste("More than two loops of the mixture modeling algorithm failed for the", dist, "distribution with", ic, "component(s). Consider choosing a lower number of components (the default is 5) via the maxcp option of the fitloops function."))
    }
  }  
}

#-----------------------------------------------
#[IC2]  bestfits
#-----------------------------------------------
#Determines the optimal number of distributions and type of distributions (as determined by BIC), was well as the other indicated functions. Also creates table of BICs which can be fed into the bicgraph function to visualize which combination of distributions and component numbers is optimal by BIC.

bestfits<-function(fitres){

  maxcp<-fitres$maxcp
  loops<-fitres$loops
  nobj<-vector("list")

  for (i in 1:maxcp){
    for (j in 1:loops){
      nobj$no[i]<-fitres$no[[i]]$bic
      nobj$sn[i]<-fitres$sn[[i]]$bic
    }
  }

  nobj$minind<-data.frame(no=as.numeric(c(NA)), sn=as.numeric(c(NA)))

  if (is.na(nobj$no[1])==T){
    nobj$minind$no<-1
  } else {
    nobj$minind$no <- (which.min(nobj$no))
  }

  if (is.na(nobj$sn[1])==T){
    nobj$minind$sn<-1
  } else {
    nobj$minind$sn <- (which.min(nobj$sn))
  }

  nobj$min<-data.frame(type=c("no", "sn"), bic=as.numeric(c(NA, NA)))

  nobj$min$type<-as.character(c("no", "sn"))
  type<-c("no", "sn")
  nobj$min$type<-type

  nobj$min[1,2]<-fitres$no[[nobj$minind$no[1]]]$bic
  nobj$min[2,2]<-fitres$sn[[nobj$minind$sn[1]]]$bic

  nobj$distnum<-which.min(nobj$min$bic)
  nobj$dist<-nobj$min$type[nobj$distnum]

  nobj$best<-fitres[[nobj$dist]][[nobj$minind[nobj$dist][1,1]]]

  if (nobj$dist=="no"){
    nobj$distp="Normal"
    nobj$distf="Normal"
  } else if (nobj$dist=="sn") {
    nobj$distp="Skew normal"
    nobj$distf="Skew.normal"
  } else {
    nobj$distp=""
  }

  model<-vector("list")
  model$sn2<-fitres$sn[[2]]
  model$snbest<-fitres$sn[[nobj$minind$sn[1]]]
  model$norm2<-fitres$no[[2]]
  model$normbest<-fitres$no[[nobj$minind$no[1]]]
  model$best<-nobj$best
  model$bictab<-data.frame(normal=c(NA)[rep(c(1), times=maxcp)], skew.normal=c(NA)[rep(c(1), times=maxcp)])
  model$bictab$normal<-nobj$no
  model$bictab$skew.normal<-nobj$sn
  componentnames<-vector()
  for (i in 1:maxcp){
    componentnames[i]<-paste(i, " Component(s):  ", sep="")  
  }    
  rownames(model$bictab)<-componentnames
  colnames(model$bictab)<-c("Normal", "Skew normal")
  model$summary<-data.frame(desc=c(paste(nobj$distp, "with", nobj$minind[nobj$dist][1,1], "components ", sep=" "), paste("Skew normal with", nobj$minind$sn, "components ", sep=" "), paste("Normal with", nobj$minind$no, "components ", sep=" "), "Skew normal with 2 components ", "Normal with 2 components "), bic=c(nobj$best$bic, nobj$sn[nobj$minind$sn], nobj$no[nobj$minind$no], nobj$sn[2], nobj$no[2]))
  rownames(model$summary)<-c("Best Overall", "Best Skew normal  ", "Best Normal", "Two Skew normal", "Two Normal")
  colnames(model$summary)<-c("Description         ", "BIC  ")

  model$bestdesc<-list(ncomp=nobj$minind[nobj$dist][1,1], dist=nobj$distf)
  model$datawithids<-fitres$datawithids

  print(paste ("The best model by BIC is", nobj$distp, "with", nobj$minind[nobj$dist][1,1], "components", sep = " "))
  print("")
  print("Bayesian Information Criterion (BIC) Matrix")
  print("BIC should be minimized and a difference of 10 BIC indicates strong evidence that the model with lower BIC is superior")
  print(model$bictab)
  print("Below is a table of the BIC's of the most common distribution and number of component combinations to base a cutpoint on")
  print(model$summary)

  return(model)

}

#-----------------------------------------------
#[IC3]  bicgraph
#-----------------------------------------------
#Plots BIC to show which combination of distributions and number of components is optimal by BIC. BIC is a criterion which should be minimized (ie we are looking for the most negative value), and a decrease in BIC of 10 or more is "strong evidence" that the model with lower BIC is superior. However in some cases you may want to choose a different distribution and this graph allows for an easy compairison of relative BIC.

bicgraph<-function(model,title="BIC by type and number of distributions",setcolor=c("magenta", "blue", "black")){
  color<-setcolor
  if (is.na(model$bictab$Normal[1])==F) {
    plot(model$bictab$Normal, pch=1, col=color[1], main=title, xlab="Number of distributions", ylab="BIC", type="o",  xaxt="n")
    lines(model$bictab$`Skew normal`, type="o", pch=2, lty=2, col=color[2])
  } else {
    plot(model$bictab$Normal, type="o", pch=2, lty=2, col=color[2], main=title, xlab="Number of distributions", ylab="BIC",  xaxt="n")
    lines(model$bictab$`Skew normal`, pch=1, col=color[1], type="o")    
  }  
  points(model$bestdesc$ncomp,model$best$bic,type="o", pch="O", col=color[3],cex=2)
  axis(1, at = seq(1, length(model$bictab$Normal), by = 1))
  suppressWarnings(legend("topright", col=c(color[1], color[2], color[3]), lty=c(1,2,0), pch = c(1, 2, 1), legend=c("Normal", "Skew normal", "Best by BIC"), cex=c(1,1,1), xjust=1, seg.len=3, title="Distribution"))

}

#-----------------------------------------------
#[IC4]  fitpick
#-----------------------------------------------
#Selects a specific distribution and format so that it can be read into the uncertainty functions. Default picks the optimized BIC function from bestfits, but this can be adjusted if the user chooses otherwise.

fitpick<-function(fitres,bestfits,dist="",ncomp=NA){
  if (dist=="" & is.na(ncomp)==T) {
    dist=bestfits$bestdesc$dist
    ncomp=bestfits$bestdesc$ncomp
  }
  if (ncomp==1) {
    stop("Single component specified so there is no cut-point")
  }

  data=fitres$datawithids$data
  n=ncomp
  if (dist!="Normal" & dist!="Skew.normal") {
    stop("Distribution must be specified as either 'Normal' or 'Skew.normal'")
  } else if ((dist=="Normal" & is.na(fitres$no[[1]]$bic)==T)| (dist=="Skew.normal" & is.na(fitres$sn[[1]]$bic)==T) | ncomp>length(fitres$no) ){
    stop("Must choose a distribution and number of components that were fitted in the fitloops function")
  } else {
    if (dist=="Normal"){
      dist="no"
    } else {
      dist="sn"
    }

    if (dist=="no"){
      distp="Normal"
    } else if (dist=="sn") {
      distp="Skew Normal"
    } else {
      distp=""
    }

    singlefit<-vector("list")
    singlefit<-fitres[[dist]][[ncomp]]

    comp<-data.frame(ogroup=singlefit$group, data=data)
    comp$index <- as.numeric(row.names(comp))
    comp <- comp[order(comp$data),] 
    comp$diff<-c(0, diff(comp$ogroup))
    nl=as.numeric(dim(comp)[1])

    #Will recode the distributions so that they are in order with 1 being the lowest
    comp$c<-vector(mode="numeric", length=nl)
    comp$c[1]<-1
    for (i in 2:nl){
      if (comp$diff[i]==0) {
        comp$c[i]=comp$c[i-1]
      } else {
        comp$c[i]=comp$c[i-1]+1
      }
    }
    comp <- comp[order(comp$index),]
    singlefit$group2<-comp$c
    comp <- comp[order(comp$data),]

    intDistErr = function() {
      tryCatch(data.frame(c=unique(comp$c), o=unique(comp$ogroup)),
               error = function(e) {data.frame(c=c(NA)[rep(c(1), times=length(comp$c))], o=c(NA)[rep(c(1), times=length(comp$c))])}) 
    }

    cross<-intDistErr()
    if (suppressWarnings(is.na(cross[1,1])==T)){
      stop("This combination of distribution and number of components yeilds interrupting distributions, which make determining a cut-point impossible. Please choose a different combination of distribution and number of components")
    }

    for (i in 1:n){
      singlefit$mu[i]<-fitres[[dist]][[ncomp]]$mu[cross$o[i]]
      singlefit$sigma2[i]<-fitres[[dist]][[ncomp]]$sigma2[cross$o[i]]
      singlefit$shape[i]<-fitres[[dist]][[ncomp]]$shape[cross$o[i]]
      singlefit$pii[i]<-fitres[[dist]][[ncomp]]$pii[cross$o[i]]
      singlefit$mu[i]<-fitres[[dist]][[ncomp]]$mu[cross$o[i]]
    }

    singlefit$obs.prob<-NULL

    #Detect if there are any interupting distributions (can also be recoded to kick someone out and tell them to pick something else:
    if (max(comp$c)>max(comp$ogroup)) {
      print("Warning: there is at least one interupting distribution, which may make the determination of a cutoff and the calculation of an indeterminate range difficult in later functions. If later functions fail, try a different combination of distribution and number of components (ideally the next most optimal by BIC).")
    }

    par<-vector("list")
    par<-data.frame(mean=c(NA)[rep(c(1), times=n)], sd=c(NA)[rep(c(1), times=n)], shape=c(NA)[rep(c(1), times=n)], pii=c(NA)[rep(c(1), times=n)])
    for (i in 1:n){
      par$mean[i]=singlefit$mu[i]
      par$sd[i]=sqrt(singlefit$sigma2[i])
      par$sigma2[i]=singlefit$sigma2[i]
      par$shape[i]=singlefit$shape[i]
      par$pii[i]=singlefit$pii[i]
    }

    parameters<-vector("list")
    parameters$singlefit<-singlefit
    parameters$par<-par
    parameters$datawithids<-fitres$datawithids
    parameters$desc<-vector("list")
    parameters$desc$ncomp<-ncomp
    parameters$desc$dist<-distp
  }

  return(parameters)
}

#-----------------------------------------------
#[IC5]  uncert
#-----------------------------------------------
#Performs uncertainty calculations.

uncert<-function(fitpickobj){

  data<-fitpickobj$datawithids$data
  par<-fitpickobj$par
  fit<-fitpickobj$singlefit
  n=fitpickobj$desc$ncomp
  #Create Uncertainties
  v<-c(seq(min(data),max(data),0.0001))

  nl<-length(v)
  PD<-vector(mode="numeric", length=n)
  PD[1]<-1
  g1<-0
  g2<-0
  Pv.D<-matrix(0, nl, n)
  Pv.DPD<-matrix(0, nl, n)
  PDT<-matrix(0, nl, 1)
  PD.v<-matrix(0, nl, n)
  for (i in 2:n) {
    PD[i]<-par$pii[i]
    PD[1]<-PD[1]-PD[i]
  }
  for (i in 1:n){
    if (fitpickobj$desc$dist=="Normal"){
      Pv.D[,i]<-dnorm(v,mean=par$mean[i], sd=par$sd[i])
    } else {
      Pv.D[,i]<-dsn(v,xi=par$mean[i],omega=par$sd[i],alpha=par$shape[i])
    }
  }
  for (i in 1:n){
    for (j in 1:nl){
      Pv.DPD[j,i]<-Pv.D[j,i]*PD[i]
    }
  }
  PDT<-rowSums (Pv.DPD)
  for (i in 1:n){
    for (j in 1:nl){
      PD.v[j,i] <- Pv.DPD[j,i] / (PDT[j])
    }
  }

  mcluncert<-1-apply(PD.v, 1, max) 
  uncertmat<-PD.v
  restrmat<-cbind(PD.v, v)
  uncertout<-vector("list")
  uncertout$uncertmat<-uncertmat
  uncertout$mcluncert<-mcluncert
  uncertout$restrmat<-restrmat
  uncertout$singlefit<-fit
  uncertout$v<-v
  uncertout$datawithids<-fitpickobj$datawithids
  uncertout$desc<-fitpickobj$desc
  return(uncertout) 
}

#-----------------------------------------------
#[IC6]  multcut
#-----------------------------------------------
#Calculates raw percentages and counts positive and negative for all possible cut points between distributions
multcut<-function(uncertobj){

  PD.v<-uncertobj$uncertmat
  data<-uncertobj$datawithids$data
  n=uncertobj$desc$ncomp
  ncomp=n
  closest<-function(uncertdf,min,max,level){
    uncertdf2 <- uncertdf[ which(uncertdf$v>=min & uncertdf$v<max), ]
    diff.cut<-abs(uncertdf2$uncertainty.v - level)
    min.diffc<-min(diff.cut)
    y<-which(min.diffc==diff.cut)
    #take first one
    z<-min(y)
    cut<-uncertdf2$v[z]  
    return(cut) 
  }
  g1<-vector("list")
  g2<-vector("list")
  uncertdf<-vector("list")
  uncertainty.v<-vector("list")
  cutpoint<-c(0)[rep(c(1), times=ncomp-1)]
  for (j in 1:(ncomp-1)){
    g1[[j]]<-0
    g2[[j]]<-0
    for (i in 1:j){
      g1[[j]]<-g1[[j]]+PD.v[,i]
    }
    for (i in (j+1):n){
      g2[[j]]<-g2[[j]]+PD.v[,i]
    }
    uncertainty.v[[j]]<-1-pmax(g1[[j]], g2[[j]])
  }

  for (j in 1:(ncomp-1)){
    uncertdf[[j]]<-data.frame(v=uncertobj$v, uncertainty.v=uncertainty.v[[j]])
    cutpoint[j]<-closest(uncertdf[[j]],min(data),max(data),1.0)
  }
  class2<-vector("list")
  for (j in 1:(ncomp-1)){
    class2[[j]]<-ifelse(data>=cutpoint[j], "positive", "negative")
  }
  class<- data.frame(matrix(unlist(class2), nrow=uncertobj$singlefit$n, byrow=F),stringsAsFactors=TRUE)
  class$data<-data

  table<-vector("list", ncomp-1)
  for (i in 1:(ncomp-1)){
    table[[i]]<-table(class2[[i]])

  }
  npos<-vector("numeric")
  nneg<-vector("numeric")
  dpos<-vector("numeric")
  dneg<-vector("numeric")
  ppos<-vector("numeric")
  pneg<-vector("numeric")
  grpname<-vector("numeric")
  for (i in 1:(ncomp-1)){
    npos[i]<-table[[i]][2]
    nneg[i]<-table[[i]][1]
    dpos[i]<-npos[i]/(npos[i]+nneg[i])
    dneg[i]<-1-dpos[i]
    ppos[i]<-paste(round(100*dpos[i], 2), "%", sep="")
    pneg[i]<-paste(round(100*dneg[i], 2), "%", sep="")
    grpname[i]<-paste("Cut between components ", i, " and ", i+1, sep="")
  }
  classtab<-data.frame( nneg, npos, dneg, dpos)
  outtab<-data.frame(nneg, npos, pneg, ppos)
  rownames(outtab) <- grpname
  colnames(outtab) <-c("Number Negative", "Number Positive", "Percent Negative", "Percent Positive")

  cutobj<-vector("list")
  cutobj$cutpoint<-cutpoint
  cutobj$class<-class
  cutobj$uncertdf<-uncertdf
  cutobj$type<-"Standard"
  cutobj$g1<-g1
  cutobj$g2<-g2
  cutobj$classtab<-classtab
  cutobj$outtab<-outtab
  cutobj$cutnames<-grpname
  #    cutobj$uncertmat<-uncertobj$uncertmat
  #    cutobj$mcluncert<-uncertobj$mcluncert
  #    cutobj$restrmat<-uncertobj$restrmat
  #    cutobj$singlefit<-uncertobj$fit
  #    cutobj$data<-uncertobj$data
  #    cutobj$desc<-uncertobj$desc
  print(outtab)

  return(cutobj)
}

#-----------------------------------------------
#[IC7]  rawuncertgraph
#-----------------------------------------------
#Displays the overall uncertainty function (which is 1-max(probability of membership to component i)) with the possible cut-points overlaid. Note: Sometimes the cutpoint does not line up exactly with a peak in the uncertainty function, which looks like a flaw, but is not. In a two component setting, the cutpoint will always occur at the peak of the uncertainty but this is not always the case when there are more than two components.
rawuncertgraph<-function(uncertobj,multcutobj,xlab="X Value",xlim=c(NA,NA),suppresslegend=F,setcolor=c(
  "green4",          #1
  "turquoise3",      #2
  "royalblue2",      #3
  "darkorchid2",     #4
  "firebrick3",      #5
  "darkorange2",     #6
  "white",           #7
  "grey",            #8
  "black"            #9
)){

  grobj<-list()
  if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
    grobj$minval<-(1/10)*trunc(10*min(uncertobj$datawithids$data)) 
    grobj$maxval<-max(uncertobj$datawithids$data)
  } else {
    grobj$minval<-xlim[1]
    grobj$maxval<-xlim[2]
  }

  plot(uncertobj$v, uncertobj$mcluncert, main=paste("Uncertainty Plot with Possible Cutpoints for", uncertobj$desc$dist, "with", uncertobj$desc$ncomp, "components", sep = " "), type="l", ylim=c(0,1), xlim=c(grobj$minval,grobj$maxval), xlab=xlab, ylab="Uncertainty")

  col<-setcolor
  ncomp<-uncertobj$desc$ncomp
  for (i in 1:(uncertobj$desc$ncomp-1)){
    abline(v =multcutobj$cutpoint[i], untf = T, col=col[i], lwd=2)
  }

  if (suppresslegend==F) {
    legtxt<-vector("character", ncomp)
    colors<-vector("character", ncomp)
    for (i in 2:ncomp){
      legtxt[i]<-paste(multcutobj$cutnames[i-1], " , x=", round(multcutobj$cutpoint[i-1], 2), sep="")
      colors[i]<-col[i-1]
    }  
    legtxt[1]<-"Uncertainty"
    colors[1]<-col[length(col)]
    linewidth<-c(2)[rep(c(1), times=ncomp)]
    legend("topright", col=colors, lwd=c(1,linewidth), legend=legtxt, xjust=1, seg.len=3)
  }
}

#-----------------------------------------------
#[IC8]  rawdistgraph
#-----------------------------------------------
#Displays the distributions of all the components for the chosen combination of distribution and number of components. 
rawdistgraph<-function(pickobj,multcutobj,xlim=c(NA,NA),xlab="Optical Density",setbreaks=100,suppresslegend=F,setcolor=c(
  "green4",          #1
  "turquoise3",      #2
  "royalblue2",      #3
  "darkorchid2",     #4
  "firebrick3",      #5
  "darkorange2",     #6
  "white",           #7
  "grey",            #8
  "black"            #9
)){
  #Parameter specification
  ncomp=pickobj$desc$ncomp
  grobj<-vector("list")
  grobj$gpar<-vector("list")
  data<-pickobj$datawithids$data
  color<-setcolor

  if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
    grobj$minval<-(1/10)*trunc(10*min(pickobj$datawithids$data)) 
    grobj$maxval<-max(data)
  } else {
    grobj$minval<-xlim[1]
    grobj$maxval<-xlim[2]
  }

  grobj$t<-seq(grobj$minval,grobj$maxval,length=1000)

  grobj$gpar<-data.frame(mean=c(NA)[rep(c(1), times=ncomp)], sd=c(NA)[rep(c(1), times=ncomp)], shape=c(NA)[rep(c(1), times=ncomp)], pii=c(NA)[rep(c(1), times=ncomp)])
  grobj$gt<-vector("list", ncomp+1)
  grobj$maxdens<-vector("numeric", ncomp)

  grobj$coloropt<-vector("list", 5)
  grobj$coloropt[[1]]<-c(color[1], color[9])
  grobj$coloropt[[2]]<-c(color[1], color[5], color[9])
  grobj$coloropt[[3]]<-c(color[1], color[3], color[5], color[9])
  grobj$coloropt[[4]]<-c(color[1], color[3], color[4], color[5], color[9])
  grobj$coloropt[[5]]<-c(color[1], color[2], color[3], color[4], color[5], color[9])
  grobj$colors<-grobj$coloropt[[ncomp]]

  for (i in 1:ncomp){
    grobj$gpar$mean[i]=pickobj$par$mean[i]
    grobj$gpar$sd[i]=sqrt(pickobj$par$sigma2[i])
    grobj$gpar$shape[i]=pickobj$par$shape[i]
    grobj$gpar$pii[i]=pickobj$par$pii[i]
  }
  #Creation of points for curves of Graph at full scale (components scaled to add to overall density)
  for (i in 1:ncomp) {
    grobj$gt[[i]] <- grobj$gpar$pii[i]*dsn(grobj$t,xi=grobj$gpar$mean[i],omega=grobj$gpar$sd[i],alpha=grobj$gpar$shape[i])
  }
  grobj$gt[[ncomp+1]]<-vector("numeric", 1000)
  for (i in 1:ncomp) {
    grobj$gt[[ncomp+1]]<-grobj$gt[[ncomp+1]]+grobj$gt[[i]]
  }
  grobj$ylim<-c(0, max(grobj$gt[[ncomp+1]], max(hist(data,setbreaks,plot=F)$density,na.rm=TRUE) ))
  grobj$xlim<-c(grobj$minval,grobj$maxval)

  dist<-c(0)[rep(c(0), times=ncomp)]
  for (i in 1:ncomp){
    if (pickobj$singlefit$shape[i]>0) {
      dist[i]<-1
    }
  }

  #Plot Graph

  hist(data, freq=F, breaks=setbreaks, xlab=xlab, ylab="Density", main=paste("Distribution of", ncomp, pickobj$desc$dist,"Components", sep = " "), border="grey", ylim=grobj$ylim, xlim=grobj$xlim)
  for (i in 1:ncomp) {
    lines(grobj$t, grobj$gt[[i]], lwd=2, col=grobj$colors[i])
  }
  lines(grobj$t, grobj$gt[[length(grobj$gt)]], lwd=2, col=grobj$colors[length(grobj$colors)], lty=2)
  legtxt<-vector("character", ncomp+1)
  colors<-vector("character", ncomp+1)
  for (i in 1:ncomp){
    legtxt[i]<-paste("Distribution of Component ", i, sep="")
    colors[i]<-grobj$colors[i]
  }  
  legtxt[ncomp+1]<-"Overall Distribution"
  colors[ncomp+1]<-grobj$colors[length(grobj$colors)]
  linetype<-c(1)[rep(c(1), times=ncomp)]
  if (suppresslegend==F) {
    legend("topright", col=colors, lwd=2, legend=legtxt, xjust=1, seg.len=3, title="Distributions", lty=c(linetype,2))
  }

  #return(grobj)
}

#-----------------------------------------------
#[IC9]  rawhistcuts
#-----------------------------------------------
#Displays which component each bar of the histogram would be assigned to.

rawhistcuts<-function(uncertobj,multcutobj,xlab="Optical Density",xlim=c(NA,NA),setbreaks=250,suppresslegend=F,setcolor=c(
  "green4",          #1
  "turquoise3",      #2
  "royalblue2",      #3
  "darkorchid2",     #4
  "firebrick3",      #5
  "darkorange2",     #6
  "white",           #7
  "grey",            #8
  "black"            #9
)){
  data<-uncertobj$datawithids$data
  grobj<-list()

  if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
    grobj$minval<-(1/10)*trunc(10*min(data)) 
    grobj$maxval<-max(data)
    hist<-hist(data, breaks=setbreaks, plot=F)
  } else {
    grobj$minval<-xlim[1]
    grobj$maxval<-xlim[2]
    datalim<-data[ which(data>=grobj$minval & data<=grobj$maxval)]
    hist<-hist(datalim, breaks=setbreaks, plot=F)
  }

  ncomp<-uncertobj$desc$ncomp
  cuts <- cut(hist$breaks, c(-Inf,multcutobj$cutpoint,Inf))

  color<-setcolor
  coloropt<-vector("list", 5)
  coloropt[[1]]<-c(color[1], color[9])
  coloropt[[2]]<-c(color[1], color[5], color[9])
  coloropt[[3]]<-c(color[1], color[3], color[5], color[9])
  coloropt[[4]]<-c(color[1], color[3], color[4], color[5], color[9])
  coloropt[[5]]<-c(color[1], color[2], color[3], color[4], color[5], color[9])
  col<-coloropt[[ncomp]]

  cuts<-as.numeric(cuts)
  colors<-vector("character", length(cuts))
  for (i in 1:max(cuts)){
    for (j in 1:length(cuts)){
      if (cuts[j]==i) {
        colors[j]<-col[i]
      }
    }
  }
  plot(hist, col=colors, xlab=xlab, ylab="Density", main=paste("Classification by", uncertobj$desc$dist, "with", ncomp, "components", sep = " "), freq=F, lty="blank")

  if (suppresslegend==F){
    legtxt<-vector("character", ncomp)
    colors<-vector("character", ncomp)
    for (i in 1:ncomp){
      legtxt[i]<-paste("Component ", i, sep="")
      colors[i]<-col[i]
    }  
    linetype<-c(1)[rep(c(1), times=ncomp)]
    legend("topright", col=colors, lwd=3, legend=legtxt, xjust=1, seg.len=3, title="Components", lty=linetype)
  }
}

#-----------------------------------------------
#[IC10] standindet
#-----------------------------------------------
#After deciding which distribution to cut after, this function provides cutpoint and standard indeterminate ranges (90 and 95), as well as classification into negative, indeterminate and positive components. It should be noted that the function is written so that it will still run if a lower or upper bound is not found, in which case all subjects below the upper bound (if lower bound not found) or above the lower bound (if upper bound not found) will be classified as indeterminate. In this case it is reccommeded that the non-standard cuts function be used to find a more descriptive indeterminate range.
standindet<-function(uncertobj,multcutobj,cutcomp=0){
  data<-uncertobj$datawithids$data
  n=uncertobj$desc$ncomp
  if (n==2){
    cutcomp=1
  } else if (cutcomp==0) {
    stop("Component to cut after must be specified. For example if you wanted to cut between components 2 and 3, set cutcomp=2")
  }
  uncertdf<-multcutobj$uncertdf[[cutcomp]]
  closest<-function(uncertdf,min,max,level){
    uncertdf2 <- uncertdf[ which(uncertdf$v>=min & uncertdf$v<max), ]
    diff.cut<-abs(uncertdf2$uncertainty.v - level)
    min.diffc<-min(diff.cut)
    y<-which(min.diffc==diff.cut)
    #take first one
    z<-min(y)
    cut<-uncertdf2$v[z]  
    return(cut) 
  }
  crosses<-function(uncertdf,min,max,level){
    uncertdf2 <- uncertdf[ which(uncertdf$v>=min & uncertdf$v<max), ]
    diff.cut<-abs(uncertdf2$uncertainty.v - level)
    min.diffc<-min(diff.cut)
    if (min.diffc<=0.001){
      y<-which(min.diffc==diff.cut)
      #take first one
      z<-min(y)
      cut<-uncertdf2$v[z]  
    } else {
      cut<-NA
    }
    return(cut)
  }
  cutpoint<-multcutobj$cutpoint[cutcomp]  
  minunclb<-closest(uncertdf,min(data)+0.0001,cutpoint,0.0001)
  minuncub<-closest(uncertdf,cutpoint,max(data),0)
  lb90<-crosses(uncertdf,minunclb,cutpoint,0.1)  
  ub90<-crosses(uncertdf,cutpoint,minuncub,0.1)
  lb95<-crosses(uncertdf,minunclb,cutpoint,0.05)  
  ub95<-crosses(uncertdf,cutpoint,minuncub,0.05)

  bound90<-c(lb90, ub90)
  bound95<-c(lb95, ub95)

  if (is.na(lb90)==T){
    message("Lower bound of the 90% indeterminate range not found because uncertainty is never below 10% to the left of the cutpoint. Consider using non-standard bound function to find more descriptive indeterminate range")
  }
  if (is.na(ub90)==T){
    message("Upper bound of the 90% indeterminate range not found because uncertainty is never below 10% to the right of the cutpoint. Consider using non-standard bound function to find more descriptive indeterminate range")
  }
  if (is.na(lb95)==T){
    message("Lower bound of the 95% indeterminate range not found because uncertainty is never below 5% to the left of the cutpoint. Consider using non-standard bound function to find more descriptive indeterminate range")
  }
  if (is.na(ub95)==T){
    message("Upper bound of the 95% indeterminate range not found because uncertainty is never below 5% to the right of the cutpoint. Consider using non-standard bound function to find more descriptive indeterminate range")
  }

  class<-data.frame(id=uncertobj$datawithids$id, data=uncertobj$datawithids$data)
  class$groupdi<-ifelse(class$data>=cutpoint, "positive", "negative")
  class$group95<-"indeterminate"
  class$group95[class$data<=bound95[1]]<-"negative"
  class$group95[class$data>=bound95[2]]<-"positive"
  class$group90<-"indeterminate"
  class$group90[class$data<=bound90[1]]<-"negative"
  class$group90[class$data>=bound90[2]]<-"positive"

  class$groupdi<-as.factor(class$groupdi)
  class$group95<-as.factor(class$group95)
  class$group90<-as.factor(class$group90)

  class$group90 = factor(class$group90,levels(class$group90)[c(2,3,1)])
  class$group95 = factor(class$group95,levels(class$group95)[c(2,3,1)])

  cutobj<-vector("list")
  cutobj$cutpoint<-cutpoint
  cutobj$bound90<-bound90
  cutobj$bound95<-bound95
  cutobj$class<-class
  cutobj$uncertdf<-uncertdf
  cutobj$type<-"Standard"
  cutobj$uncertobj<-uncertobj$uncertmat
  cutobj$density<-vector("list")
  cutobj$density$g1<-multcutobj$g1[[cutcomp]]
  cutobj$density$g2<-multcutobj$g2[[cutcomp]]
  cutobj$desc<-uncertobj$desc
  cutobj$desc$cutcomp<-cutcomp
  cutobj$datawithids<-uncertobj$datawithids

  return(cutobj) 
}

#-----------------------------------------------
#[IC11]  specindet
#-----------------------------------------------
#Provides cutpoint and allows the user to specify a maximum tolerable uncertianty which will be used to calculate an indeterminate ranges, as well as classification into negative, indeterminate and positive components. Maximum tolerable uncertianty is 1-certainty level (ex: uncertlevel=0.05 for 95% indetermintate range since the 95% certainty means that uncertainty of classification is less than 0.05 for all samples classified as positive or negative).
specindet<-function(uncertobj,multcutobj,certlevel,cutcomp=0){
  uncertlevel<-1-certlevel
  uncertlevelpct<-certlevel

  PD.v<-uncertobj$uncertmat
  data<-uncertobj$datawithids$data
  n=uncertobj$desc$ncomp

  if (n==2){
    cutcomp=1
  } else if (cutcomp==0) {
    stop("Component to cut after must be specified. For example if you wanted to cut between components 2 and 3, set cutcomp=2")
  }
  if (certlevel>=1 | certlevel<0.5){
    stop("Certainty level must be between 0.5 and 1, for example if you wanted a certainty level of 90% set certlevel=0.90")
  }

  uncertdf<-multcutobj$uncertdf[[cutcomp]]
  closest<-function(uncertdf,min,max,level){
    uncertdf2 <- uncertdf[ which(uncertdf$v>=min & uncertdf$v<max), ]
    diff.cut<-abs(uncertdf2$uncertainty.v - level)
    min.diffc<-min(diff.cut)
    y<-which(min.diffc==diff.cut)
    z<-min(y)
    cut<-uncertdf2$v[z]  
    return(cut) 
  }
  crosses<-function(uncertdf,min,max,level){
    uncertdf2 <- uncertdf[ which(uncertdf$v>=min & uncertdf$v<max), ]
    diff.cut<-abs(uncertdf2$uncertainty.v - level)
    min.diffc<-min(diff.cut)
    if (min.diffc<=0.001){
      y<-which(min.diffc==diff.cut)
      #take first one
      z<-min(y)
      cut<-uncertdf2$v[z]  
    } else {
      cut<-NA
    }
    return(cut)
  }
  cutpoint<-closest(uncertdf,min(data),max(data),0.5)
  minunclb<-closest(uncertdf,min(data)+0.0001,cutpoint,0.0001)
  minuncub<-closest(uncertdf,cutpoint,max(data),0)
  lbuns<-crosses(uncertdf,minunclb,cutpoint,uncertlevel)  
  ubuns<-crosses(uncertdf,cutpoint,minuncub,uncertlevel)
  bound<-c(lbuns, ubuns)

  class<-data.frame(id=uncertobj$datawithids$id, data=uncertobj$datawithids$data)
  class$groupdi<-ifelse(class$data>=cutpoint, "positive", "negative")
  class$group<-"indeterminate"
  class$group[class$data<=bound[1]]<-"negative"
  class$group[class$data>=bound[2]]<-"positive"

  class$groupdi<-as.factor(class$groupdi)
  class$group<-as.factor(class$group)
  class$group = factor(class$group,levels(class$group)[c(2,3,1)])

  cutobj<-vector("list")
  cutobj$cutpoint<-cutpoint
  cutobj$bound<-bound
  cutobj$class<-class
  cutobj$uncertdf<-uncertdf
  cutobj$type<-"Non.Standard"
  cutobj$uncertobj<-uncertobj$uncertmat
  cutobj$density<-data.frame(g1=multcutobj$g1[[cutcomp]], g2=multcutobj$g2[[cutcomp]])
  cutobj$desc<-uncertobj$desc
  cutobj$desc$cutcomp<-cutcomp
  cutobj$desc$uncertlevel<-uncertlevelpct
  cutobj$datawithids<-uncertobj$datawithids

  return(cutobj) 
}

#-----------------------------------------------
#[IC12]  cutuncertgraph
#-----------------------------------------------
#Displays the uncertainty function after components have been combined to create positive and negative components with cut-points and bounds of indeterminate range(s) overlaid. Accepts results from both standard and nonstandard cutpoints (from the functions standindet and specindet respectively).
cutuncertgraph<-function(cutobj,xlab="Optical Density",xlim=c(NA,NA),suppresslegend=F,setcolor=c(
  "green4",          #1
  "turquoise3",      #2
  "royalblue2",      #3
  "darkorchid2",     #4
  "firebrick3",      #5
  "darkorange2",     #6
  "white",           #7
  "grey",            #8
  "black"            #9
)){
  color=setcolor

  if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
    plot(cutobj$uncertdf$v, cutobj$uncertdf$uncertainty.v, main=paste("Uncertainty Plot of", cutobj$desc$dist, "with", cutobj$desc$ncomp, "components, cut between distributions", cutobj$desc$cutcomp, "and", (cutobj$desc$cutcomp+1), sep = " "),type="l", ylim=c(0,1), xlab=xlab, ylab="Uncertainty")
  } else {
    grobj$minval<-xlim[1]
    grobj$maxval<-xlim[2]
    plot(cutobj$uncertdf$v, cutobj$uncertdf$uncertainty.v, main=paste("Uncertainty Plot of", cutobj$desc$dist, "with", cutobj$desc$ncomp, "components, cut between distributions", cutobj$desc$cutcomp, "and", (cutobj$desc$cutcomp+1), sep = " "),type="l", ylim=c(0,1), xlab=xlab, ylab="Uncertainty", xlim=c(grobj$minval,grobj$maxval))
  }

  abline(v =cutobj$cutpoint, untf = T, col=color[6], lwd=1)
  if (cutobj$type=="Standard") {
    abline(h =0.05, untf = T, col=color[4], lwd=1, lty=2)
    abline(h =0.1, untf = T, col=color[2], lwd=1, lty=2)
    abline(v =cutobj$bound90, untf = T, col=color[2], lwd=1)
    abline(v =cutobj$bound95, untf = T, col=color[4], lwd=1)
    legtxt<-c("Uncertainty",
              paste("Cutpoint=", round(cutobj$cutpoint, 2), sep =""), 
              "Uncertainty=0.10",
              "Indeterminate Range with", 
              paste("90% Certainty: (", round(cutobj$bound90[1], 2), " , ", round(cutobj$bound90[2], 2), ")", sep =""),
              "Uncertainty=0.05",
              "Indeterminate Range with", 
              paste("95% Certainty: (", round(cutobj$bound95[1], 2), " , ", round(cutobj$bound95[2], 2), ")", sep ="")
    )
    if (suppresslegend==F){
      legend("topright", col=c(color[9], color[6], color[2], color[2], "white", color[4], color[4], "white"), lwd=1, lty=c(1,1,2,1,1,2,1,1),legend=legtxt, xjust=1, seg.len=1, title="Uncertainties, Cutpoint, and Boundaries")
      }
  } else {
    abline(h =(1-cutobj$desc$uncertlevel), untf = T, col=color[3], lwd=1, lty=2)
    abline(v =cutobj$bound, untf = T, col=color[3], lwd=1)
    legtxt<-c("Uncertainty",
              paste("Cutpoint=", round(cutobj$cutpoint, 2), sep =""), 
              paste("Uncertainty=", (1-cutobj$desc$uncertlevel), sep =""),
              "Indeterminate Range with", 
              paste((100*cutobj$desc$uncertlevel), "% Certainty: (", round(cutobj$bound[1], 2), " , ", round(cutobj$bound[2], 2), ")", sep ="")
    )
    if (suppresslegend==F){
      legend("topright", col=c(color[9],color[6], color[3], color[3], "white"), lwd=1, lty=c(1,1,2,1), legend=legtxt, xjust=1, seg.len=1, title="Cutpoint, Uncertainties, and Boundaries")
    }
  }
}

#-----------------------------------------------
#[IC13]  cutdistgraph
#-----------------------------------------------
#Displays the distributions of the positive and negative components with cut-points and bounds of indeterminate range(s) overlaid. Accepts results from both standard and nonstandard cutpoints (from the functions standindet and specindet respectively). 
cutdistgraph<-function(cutobj,pickobj,xlim=c(NA,NA),xlab="Optical Density",setbreaks=100,suppresslegend=F,setcolor=c(
  "green4",          #1
  "turquoise3",      #2
  "royalblue2",      #3
  "darkorchid2",     #4
  "firebrick3",      #5
  "darkorange2",     #6
  "white",           #7
  "grey",            #8
  "black"            #9
)){
  #Parameter specification
  ncomp<-cutobj$desc$ncomp
  cutcomp<-cutobj$desc$cutcomp
  cutobj<-cutobj
  data<-pickobj$datawithids$data
  color<-setcolor

  grobj<-vector("list")
  grobj$gpar<-vector("list")

  if (is.na(xlim[1])==T | is.na(xlim[2])==T) { 
    grobj$minval<-(1/10)*trunc(10*min(pickobj$datawithids$data))     
    grobj$maxval<-max(data)
  } else {
    grobj$minval<-xlim[1]
    grobj$maxval<-xlim[2]
  }

  grobj$t<-seq(grobj$minval,grobj$maxval,length=1000)

  grobj$gpar<-data.frame(mean=c(NA)[rep(c(1), times=ncomp)], sd=c(NA)[rep(c(1), times=ncomp)], shape=c(NA)[rep(c(1), times=ncomp)], pii=c(NA)[rep(c(1), times=ncomp)])
  grobj$gs<-vector("list", ncomp+1)
  grobj$gt<-vector("list", 2+1)
  grobj$maxdens<-vector("numeric", ncomp)
  grobj$colors<-c(setcolor[1], setcolor[5], setcolor[9])
  grobj$t<-cutobj$uncertdf$v

  for (i in 1:ncomp){
    grobj$gpar$mean[i]=pickobj$par$mean[i]
    grobj$gpar$sd[i]=sqrt(pickobj$par$sigma2[i])
    grobj$gpar$shape[i]=pickobj$par$shape[i]
    grobj$gpar$pii[i]=pickobj$par$pii[i]
  }
  #Creation of points for curves of Graph at full scale (components scaled to add to overall density)
  for (i in 1:ncomp) {
    grobj$gs[[i]] <- grobj$gpar$pii[i]*dsn(grobj$t,xi=grobj$gpar$mean[i],omega=grobj$gpar$sd[i],alpha=grobj$gpar$shape[i])
  }
  grobj$gt[[1]]<-0
  grobj$gt[[2]]<-0
  grobj$gt[[3]]<-0
  for (i in 1:cutcomp){
    grobj$gt[[1]]<-grobj$gt[[1]]+grobj$gs[[i]]
  }
  for (i in (cutcomp+1):ncomp){
    grobj$gt[[2]]<-grobj$gt[[2]]+grobj$gs[[i]]
  }
  for (i in 1:ncomp) {
    grobj$gt[[3]]<-grobj$gt[[3]]+grobj$gs[[i]]
  }

  grobj$ylim<-c(0, max(grobj$gt[[2+1]], max(hist(data,setbreaks,plot=F)$density,na.rm=TRUE) ))
  grobj$xlim<-c(round(min(pickobj$datawithids$data),digits=1),grobj$maxval)

  #Plot Graph
  hist(data, freq=F, breaks=setbreaks, xlab=xlab, ylab="Density", main=paste(cutobj$desc$dist, "with", cutobj$desc$ncomp, "components, cut between distributions", cutobj$desc$cutcomp, "and", (cutobj$desc$cutcomp+1), sep = " "), border=color[8], ylim=grobj$ylim, xlim=grobj$xlim)

  for (i in 1:2) {
    lines(cutobj$uncertdf$v, grobj$gt[[i]], lwd=2, col=grobj$colors[i])
  }
  lines(grobj$t, grobj$gt[[length(grobj$gt)]], lwd=2, col=grobj$colors[length(grobj$colors)], lty=2)
  abline(v =cutobj$cutpoint, untf = T, col=color[6], lwd=1)
  if (cutobj$type=="Standard") {
    abline(v =cutobj$bound90, untf = T, col=color[2], lwd=1)
    abline(v =cutobj$bound95, untf = T, col=color[4], lwd=1)
    if (suppresslegend==F) {
      legtxt<-c("Negative Distribution",
                "Positive Distribution",
                "Overall Distribution",
                paste("Cutpoint=", round(cutobj$cutpoint, 2), sep =""), 
                "Indeterminate Range with", 
                paste("90% Certainty: (", round(cutobj$bound90[1], 2), " , ", round(cutobj$bound90[2], 2), ")", sep =""),
                "Indeterminate Range with", 
                paste("95% Certainty: (", round(cutobj$bound95[1], 2), " , ", round(cutobj$bound95[2], 2), ")", sep ="")
      )
      legend("topright", col=c(color[1], color[5], color[9], color[6], color[2], "white", color[4], "white"), lwd=c(2,2,2,1,1,1,1,1), legend=legtxt, xjust=1, seg.len=3, title="Distributions, Cutpoint, and Boundaries", lty=c(1,1,2,1,1,1,1,1))
    }
  } else {
    abline(v =cutobj$bound, untf = T, col=color[3], lwd=1)
    if (suppresslegend==F){
      legtxt<-c("Negative Distribution",
                "Positive Distribution",
                "Overall Distribution",
                paste("Cutpoint=", round(cutobj$cutpoint, 2), sep =""), 
                "Indeterminate Range with", 
                paste((cutobj$desc$uncertlevel*100), "% Certainty: (", round(cutobj$bound[1], 2), " , ", round(cutobj$bound[2], 2), ")", sep ="")
      )
      legend("topright", col=c(color[1], color[5], color[9], color[6], color[3], "white"), lwd=c(2,2,2,1,1,1), legend=legtxt, xjust=1, seg.len=3, title="Distributions, Cutpoint, and Boundaries", lty=c(1,1,2,1,1,1))
    }
  }

  #return(grobj)

}

#-----------------------------------------------
#[IC14]  outputdata
#-----------------------------------------------
outputdata<-function(standindetobj=NULL,specindetobj=NULL,fileandpathname=NULL){
  if (is.null(standindetobj)==T & is.null(specindetobj)==T) {
    print("At least one post-cut object must be included")
  } else {
    if (is.null(standindetobj)==T) {
      names(specindetobj$class)<-c(names(specindetobj$class[1:3]),paste('group', (100*specindetobj$desc$uncertlevel), sep=''))
      dataout<-specindetobj$class
    } else if (is.null(specindetobj)==T){
      dataout<-standindetobj$class
    } else {
      datastand<-standindetobj$class[c("id", "group90", "group95")]
      names(specindetobj$class)<-c(names(specindetobj$class[1:3]),paste('group', (100*specindetobj$desc$uncertlevel), sep=''))
      datanonst<-specindetobj$class
      dataout<-merge(datanonst,datastand,by.x="id", by.y="id")
    }
    if (is.null(fileandpathname)==F){
      write.csv(dataout, file =paste(fileandpathname, ".csv", sep=""), row.names=F)
    }
  }
  return(dataout)
}

#-----------------------------------------------
#[IC15]  summarytable
#-----------------------------------------------
#Creates table which summarizes the results of the cutting functions, yeilding the cutpoint, indeterminate ranges, counts and percentages for the dichotomous cut, 90 and 95 indeterminate ranges and (if desired) one non-standard uncertainty level.
summarytable<-function(outdataobj,standindetobj=NULL,specindetobj=NULL){
  comptable <- function(table){
    dftable<-as.data.frame(table)
    len<-nrow(dftable)
    dftable$orders<-c(NA)[rep(c(1), times=len)]
    for (i in 1:len){
      if (as.character(dftable$Var1[i])=="negative"){
        dftable$orders[i]<-0
      } else if (as.character(dftable$Var1[i])=="indeterminate"){
        dftable$orders[i]<-1
      } else if (as.character(dftable$Var1[i])=="positive"){
        dftable$orders[i]<-2
      }
    }
    row.names(dftable)<-dftable$Var1
    dftable$Var1<-NULL
    if (is.na(dftable["indeterminate", 1])==T) {
      Freq<-c(dftable$Freq,0)
      grps2<-as.data.frame(Freq)
      x<-c(row.names(dftable),"indeterminate")
      row.names(grps2)<-x
      grps2$orders<-c(dftable$orders,1)
      dftable<-grps2
      Freq<-NULL
      x<-NULL
    }
    if (is.na(dftable["negative", 1])==T) {
      Freq<-c(dftable$Freq,0)
      grps2<-as.data.frame(Freq)
      x<-c(row.names(dftable),"negative")
      row.names(grps2)<-x
      grps2$orders<-c(dftable$orders,0)
      dftable<-grps2
      Freq<-NULL
      x<-NULL
    }
    if (is.na(dftable["positive", 1])==T) {
      Freq<-c(dftable$Freq,0)
      grps2<-as.data.frame(Freq)
      x<-c(row.names(dftable),"positive")
      row.names(grps2)<-x
      grps2$orders<-c(dftable$orders,2)
      dftable<-grps2    
      Freq<-NULL
      x<-NULL
    }
    dftable<-dftable[order(dftable$orders),]
    return(dftable)
  }

  grps<-table(outdataobj$groupdi)
  grps<-comptable(grps)
  grps$group<-row.names(grps)
  grps$groupdi<-grps$Freq
  grps$Freq<-NULL

  if (ncol(outdataobj)!=5 ){
    grpsns<-table(outdataobj[,4])
    grpsns<-comptable(grpsns)
    grps$gns<-grpsns$Freq
    names(grps)<-c(names(grps[1:3]),paste(names(outdataobj)[4]))
  }

  if (ncol(outdataobj)!=4 ){ 
    grps90<-table(outdataobj$group90)
    grps90<-comptable(grps90)
    grps$group90<-grps90$Freq
    grps95<-table(outdataobj$group95)
    grps95<-comptable(grps95)
    grps$group95<-grps95$Freq
  }

  names(grps)

  for (i in names(grps)){
    nneg<-vector("numeric")
    nind<-vector("numeric")
    npos<-vector("numeric")

    dneg<-vector("numeric")  
    dind<-vector("numeric")
    dpos<-vector("numeric")

    pneg<-vector("numeric")
    pind<-vector("numeric")
    ppos<-vector("numeric")
  }

  vars<-c("groupdi", "group90", "group95", "group80")

  grpname<-vector("numeric")
  for (i in 1:(ncol(grps)-2)){
    nneg[i]<-grps[1,i+2]
    nind[i]<-grps[2,i+2] 
    npos[i]<-grps[3,i+2]
  }
  for (i in 1:(ncol(grps)-2)){
    dneg[i]<-nneg[i]/(npos[i]+nind[i]+nneg[i])
    dind[i]<-nind[i]/(npos[i]+nind[i]+nneg[i])
    dpos[i]<-npos[i]/(npos[i]+nind[i]+nneg[i])
    ppos[i]<-paste(round(100*dpos[i], 2), "%", sep="")
    pind[i]<-paste(round(100*dind[i], 2), "%", sep="")
    pneg[i]<-paste(round(100*dneg[i], 2), "%", sep="")
  }
  if (ncol(grps)==4){
    grpname<-c("Raw Cutoff", paste("Cutoff with ", specindetobj$desc$uncertlevel*100, "% Certainty", sep=""))
  } else if (ncol(grps)==5) {
    grpname<-c("Raw Cutoff", "Cutoff with 90% Certainty", "Cutoff with 95% Certainty")
  } else if (ncol(grps)==6) {
    grpname<-c("Raw Cutoff", paste("Cutoff with ", specindetobj$desc$uncertlevel*100, "% Certainty", sep=""), "Cutoff with 90% Certainty", "Cutoff with 95% Certainty")
  }

  classtab<-data.frame(nneg, nind, npos, dneg, dind, dpos)
  outtab<-data.frame(nneg, nind, npos, pneg, pind, ppos)
  rownames(outtab) <- grpname
  colnames(outtab) <-c("No. Neg", "No. Indet", "No. Pos", "% Neg", "% Indet", "% Pos")

  model<-vector("list")
  if (is.null(specindetobj)==F){
    model$cutpoint<-specindetobj$cutpoint
    model$desc<-specindetobj$desc
    model$bounds<-vector("list")
    model$bounds$boundns<-specindetobj$bound
  } else {
    model$cutpoint<-standindetobj$cutpoint     
    model$desc<-standindetobj$desc
    model$bounds<-vector("list")
    model$bounds$bound90<-standindetobj$bound90
    model$bounds$bound95<-standindetobj$bound95
  }
  if (is.null(standindetobj)==F & is.null(specindetobj)==F){
    model$bounds$bound90<-standindetobj$bound90
    model$bounds$bound95<-standindetobj$bound95
  }
  model$classtab<-classtab
  model$outtab<-outtab
  model$classification<-outdataobj

  message(paste("The absolute cutoff is", round(standindetobj$cutpoint,3), sep=" "))
  message("Boundries of Indeterminates:")
  if (is.null(specindetobj)==F) {
    message(paste(specindetobj$desc$uncertlevel*100, "% Certainty Range: (", round(specindetobj$bound[1], 3), ", ", round(specindetobj$bound[2], 3), ")", sep=""))
  }
  message(paste("90% Certainty Range: (", round(standindetobj$bound90[1], 3), ", ", round(standindetobj$bound90[2], 3), ")", sep=""))
  message(paste("95% Certainty Range: (", round(standindetobj$bound95[1], 3), ", ", round(standindetobj$bound95[2], 3), ")", sep=""))

  return(model)
}

#-----------------------------------------------
#[IC16]  doallprobsfail
#-----------------------------------------------
#Performs all of the above functions, but is very unlikely to function since in the majority of cases a the two component case will not be optimal by BIC
doallprobsfail<-function(datawithids,fitres=NA){

  if (length(fitres)==1){
    fitres5<-fitloops(datawithids)
  }

  bestfit5<-bestfits(fitres5)
  bicgraph(bestfit5)
  pickobj5<-fitpick(fitres5,bestfit5)
  uncertobj5<-uncert(pickobj5)
  multcutobj5<-multcut(uncertobj5)
  rawdistgraph(pickobj5)
  rawuncertgraph(uncertobj5,multcutobj5)
  rawhistcuts(uncertobj5,multcutobj5)
  standindetpts5<-standindet(uncertobj5,multcutobj5)
  cutuncertgraph(standindetpts5)
  cutdistgraph(standindetpts5, pickobj5)
  nsindet5<-nsstandindets(uncertobj5,multcutobj5,certlevel=0.8,cutcomp=1)
  cutuncertgraph(nsindet5)
  cutdistgraph(nsindet5, pickobj5)
  outdata5<-outputdata(standindetpts5,nsindet5)
  summary<-summarytable(outdata5, standindetpts5, nsindet5)

}

#-----------------------------------------------
#[IC17]  twobestdorest
#-----------------------------------------------
#Performs all functions after bestfit if the optimal number of components is 2.
twobestdorest<-function(fitloopsobj,bestfit){

  pickobj<-fitpick(fitloopsobj,bestfit)
  uncertobj<-uncert(pickobj)
  multcutobj<-multcut(uncertobj)
  rawdistgraph(pickobj)
  rawuncertgraph(uncertobj,multcutobj)
  rawhistcuts(uncertobj,multcutobj)
  scptsobj<-standindet(uncertobj,multcutobj)
  cutuncertgraph(scptsobj)
  cutdistgraph(scptsobj, pickobj)
  outdata<-outputdata(standindetpts,nsindet)
  summary<-summarytable(outdata, standindetpts, nsindet)

  model<-vector("list")
  model$uncertobj<-uncertobj
  model$multcutobj<-multcutobj
  model$scptsobj<-scptsobj
  model$outdataobj<-outdata
  model$summaryobj<-summary

  return(model)

}

```

```{r example1}

examp1<-read.csv(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex1.csv")

hist(examp1$data, breaks=100, main="Histogram of Data")
dens<-density(examp1$data)
plot(dens, main="Density of Data")

fitloopsobject<-fitloops(datawithids=examp1)
bestfitsobject<-bestfits(fitres=fitloopsobject)
bicgraph(model=bestfitsobject)

fitpickobject<-fitpick(fitres=fitloopsobject,bestfits=bestfitsobject)

uncertobject<-uncert(fitpickobj=fitpickobject)
multcutobject<-multcut(uncertobj=uncertobject)

rawuncertgraph(uncertobj=uncertobject,multcutobj=multcutobject)
rawdistgraph(pickobj=fitpickobject)
rawhistcuts(uncertobj=uncertobject,multcutobj=multcutobject)

standindetobject<-standindet(uncertobj=uncertobject,multcutobj=multcutobject,cutcomp=1)
specindetobject<-specindet(uncertobj=uncertobject,multcutobj=multcutobject,cutcomp=1,certlevel=0.8)

cutuncertgraph(cutobj=standindetobject)
cutdistgraph(cutobj=standindetobject, pickobj=fitpickobject)
cutuncertgraph(cutobj=specindetobject)
cutdistgraph(cutobj=specindetobject, pickobj=fitpickobject)
outdataobject<-outputdata(standindet=standindetobject,specindet=specindetobject)
summaryobject<-summarytable(outdataobj=outdataobject, standindetobj=standindetobject, specindetobj=specindetobject)
outtable<-summaryobject$outtab
outtable

```

```{r example2}

examp2<-read.csv(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex2.csv", colClasses=c("NULL", NA, NA))

hist(examp2$data, breaks=100, main="Histogram of Data")
dens<-density(examp2$data)
plot(dens, main="Density of Data")

fitloopsobject<-fitloops(datawithids=examp2)
#might not run, if it doesn't, then load the full file: 
load(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex2fitloops")
fitloopsobject<-ex2
bestfitsobject<-bestfits(fitres=fitloopsobject)
bicgraph(model=bestfitsobject)

fitpickobject<-fitpick(fitres=fitloopsobject,bestfits=bestfitsobject)

uncertobject<-uncert(fitpickobj=fitpickobject)
multcutobject<-multcut(uncertobj=uncertobject)

rawuncertgraph(uncertobj=uncertobject,multcutobj=multcutobject)
rawdistgraph(pickobj=fitpickobject)
rawhistcuts(uncertobj=uncertobject,multcutobj=multcutobject)

standindetobject<-standindet(uncertobj=uncertobject,multcutobj=multcutobject,cutcomp=1)
specindetobject<-specindet(uncertobj=uncertobject,multcutobj=multcutobject,cutcomp=1,certlevel=0.8)

cutuncertgraph(cutobj=standindetobject)
cutdistgraph(cutobj=standindetobject, pickobj=fitpickobject)
cutuncertgraph(cutobj=specindetobject)
cutdistgraph(cutobj=specindetobject, pickobj=fitpickobject)
outdataobject<-outputdata(standindet=standindetobject,specindet=specindetobject)
summaryobject<-summarytable(outdataobj=outdataobject, standindetobj=standindetobject, specindetobj=specindetobject)
outtable<-summaryobject$outtab
outtable

```

```{r while loop}

stop<-F
iter<-0
comp<-0

  loopErr = function(data) {
    tryCatch(smsn.mix(data, nu =3, g = 5, get.init = TRUE, criteria = TRUE, group = TRUE, family = "Normal", calc.im=FALSE, obs.prob=TRUE, iter.max=5000, kmeans.param=list(iter.max = 100, n.start = 50)),
             error = function(e) {message(paste("Error in iteration", iter, sep=" ")); 
               NA}) 
  }

#fit5<-list()
#while (stop==F){
#  iter<-iter+1
#  set.seed(iter*10000+iter^2+iter)    
#  fit5[iter]<- loopErr(examp2$data)
#  if (is.na(fit5[iter])==F){
#    stop<-T
#  }
#}

```