### 2017-03-22 PaperEx170308.Rmd
Cutoff_Algorithm file now only contains function definitions, and the example remains but has been commented out. As such it can be called in from other documents to solely use the functions

---
title: "Mixture Modeling Paper Results"
author: "Sarah Sullivan"
date: "March 8, 2017"
output: word_document
---

```{r set_knitr_chunk_options, echo = F, warning=F, message=F}
rm(list = ls())

library(knitr)
library(markdown)
opts_chunk$set(message=FALSE, echo = F,
               warning = F,fig.height = 5, fig.width = 10, tidy = F)

#Work
#opts_knit$set(root.dir = "C:/Users/ntdintern/Google Drive/Sarah Sullivan/Code")
#School
#opts_knit$set(root.dir = "H:/TFGH/Code")
#Laptop
opts_knit$set(root.dir = "/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/")
```

```{r loadfunctions, echo=F}
source("./Cutoff_Algorithm.R")
```

#Example 1
```{r example1, echo=T}

examp1<-read.csv(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex1.csv", colClasses=c("NULL", NA, NA))

hist(examp1$data, breaks=100, main="", xlab="Optical Density")
dens1<-density(examp1$data)
plot(dens1, main="", xlab="Optical Density")

#fitloopsobject1<-fitloops(datawithids=examp2)
  #might not run due to 5 component solution not being easily optimized, if it doesn't run either load the full file, or run fitloopsobject1<-fitloops(datawithids=examp2,maxcp=4): 
load(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex1fitloops")

bicgraph(fitobj=fitloopsobject1, title="")

modelpickobject1<-modelpick(fitloopsobject1)

rawuncertgraph(modelpickobject1, title="", xlab="Optical Density")
rawdistgraph(modelpickobject1, title="", xlab="Optical Density")
rawhistcuts(modelpickobject1, title="", xlab="Optical Density")

cutoffobject1<-cutoff(modelpickobj = modelpickobject1, newcertlevel = 0.85)

cutuncertgraph(cutobj = cutoffobject1, title="", xlab="Optical Density")
cutdistgraph(cutobj = cutoffobject1, title="", xlab="Optical Density")

summaryobject1<-summaryout(cutobj = cutoffobject1)
summaryobject1$outtab

```

#Example 2

```{r example2, echo=T}

examp2<-read.csv(file="/Users/sarahsullivan/Documents/SullivanDocs/Rwd/cutoff/ex2.csv", stringsAsFactors = F, colClasses=c("NULL", NA, NA))

hist(examp2$data, breaks=100, main="", xlab="Optical Density")
dens2<-density(examp2$data)
plot(dens2, main="", xlab="Optical Density")

fitloopsobject2<-fitloops(datawithids = examp2)

bicgraph(fitobj = fitloopsobject2, title="")

modelpickobject2<-modelpick(fitobj = fitloopsobject2)

rawuncertgraph(modelpickobj = modelpickobject2, title="", xlab="Optical Density")
rawdistgraph(modelpickobj = modelpickobject2, title="", xlab="Optical Density")
rawhistcuts(modelpickobj = modelpickobject2, title="", xlab="Optical Density")

cutoffobject2<-cutoff(modelpickobj = modelpickobject2, cutcomp = 1, newcertlevel = 0.85)

cutuncertgraph(cutobj = cutoffobject2, title="", xlab="Optical Density")
cutdistgraph(cutobj = cutoffobject2, title="", xlab="Optical Density")

summaryobject2<-summaryout(cutobj = cutoffobject2)
summaryobject2$outtab
```